Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2018-05-30 hpmoon-asoco.tex
Added figures for issue #22 and addressing some @JJ 's comments

  Real world optimization problems usually have many dimensions and
  several objectives. The high dimensionality requires exceptional
  computing power, and the fact that several dimensions have to be
  searched at the same time add additional difficulty.
% One of the most extended methods to deal with high-dimensional
% multi-objective optimization problems are the distributed
% co-evolutionary algorithms.
% ----------------- This sentence is meaningless. It should be
% substituted by other stating that co-evolutionary algorithms can
% deal with it for some reason
% Besides, we're no longer saying that they are co-evolutionary in the
% title, but "island-based selective operator application". This
% should go to the abstract too - JJ

  In these methods %which ones? coevolutionary or selective operator
                   %application? - JJ
  each processor deals only with a part of the decision
space, that is, the individuals belonging to the same sub-population,
which are called {\em islands} in this context, explore the same subset of decision variables. However, only
decomposable problems could be strictly implemented in parallel this
way. A new approach, which we call {\em selective operator application} is presented here to deal with un-decomposable
problems: % so we are saying that high-dimensional multi-objective
          % problems are a challenge, but we are presenting a solution
          % for non-decomposable problems. What's what? - JJ
instead of sharing parts
of individuals, each processor applies the variation operators to a
specific subset of the whole individual. % this solves... - JJ
We have previously proved
that extending the length of these subsets, and therefore, sharing
a part of the search space among islands, implies an improvement in the
results % which results? - JJ
but in this paper, a new method to automatically
adapt the size of the overlapping, or shared sections, of the
individuals has been compared to other collaborative evolutionary
techniques considering different number of islands, chromosome sizes
and benchmarks. The analysis of the obtained experimental results, by
using different metrics, shows that our approach can provide
statistically significant improvements with respect to the base
algorithm. % in high-dimensional multi-objective problems? - JJ
Results also show that the relation between the number of
islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables.  % This
                                % solves a problem or challenge we
                                % hadn't stated. - JJ

@@

\begin{itemize}
\item Hypervolume (HV): measures the area formed by all non-dominated solutions found with respect to a reference point. Higher values imply better quality of the PF. Figure \ref{fig:hypervolume} clarifies how the hypervolume is calculated.
\item Inverted Generational distance (IGD): calculates the distance of the obtained set of solutions to the optimal PF. Therefore, this metric requires the optimal PF found in the literature, or the theoretical one. In this metric, the lower the better. %Pablo: theoretical?
\item Spread (S): Measures the spread between solutions, taking into account the euclidean distance between consecutive solutions. As in previous metric, the lower value, the better, as it implies solutions distributed along all the PF.
\end{itemize}

@@

Different metrics, previously explained, have been used
to calculate the quality of the obtained PFs in each configuration. As
some of the metrics  (such as HV) require a reference point to be
calculated, the point (1,9) has been chosen as reference, as none of
the generated PFs in all runs are dominated by it. Metrics are then
normalized respecting to that point. A Kruskal-Wallis significance
test has been performed to the metrics of all runs of the
configurations, as the Kolmogorov-Smirnov test detected non-normal
distributions. The average results for each configuration are shown in
Table \ref{tab:results512} (for 512 dimensions) and Table
\ref{tab:results2048} (for 2048). As previously explained, when $P=8$
the results of A are equivalent to the obtained with D (because
$c=0$), and when $P=16$ they are equivalent to O ($c=1$). We refer this equivalence with the word {\em Equiv-} in tables. For example, Equiv-D in an HV cell  means that the HV of A is the same than the value in D column, because the behaviour of this configuration is the same than executing D.

@@

%EXPLANATION OF THE RESULTS
This can be explained comparing the number of non-dominated solutions
of each front and the average number of generations (Table
\ref{tab:sols512} and Table \ref{tab:sols2048}). With both chromosome
lengths, increasing the number of islands implies more generations
with all the methods (logically, as there are less individuals in each
island). But also, the SLA methods
approach to the number of generations with the baseline when
increasing the number of islands. However, with $L=2048$ the migration
time is big enough  to not improve the number of generations
with respect to the baseline, even in the lower number of islands, as
it happens with $L=512$. Therefore, more generations do not
necessarily improve the solution of the global PF, but focusing on
different elements of the chromosome. As  previously stated, every
island is unaware of the other islands PFs, and they are trying to
optimize their solutions independently. With respect to the average
number of solutions per front, there is a clear difference 
 with the baseline with $L=512$, where this value is in the most
of the cases, less than a half. The number of non-dominated solutions
also implies a better Spread indicator, where the baseline 
 obtains better (or not significantly different) values in the majority of the compared configurations. %TENGO QUE COMENTAR MAS LOS RESULTADOS Y
                              %ENTENDERLO TODO Y JUSTIFICARLO
                              %FINAMENTE
%Almost always no es muy preciso - JJ Pablo: arreglao

>>>>

  \textcolor{blue}{Real world optimization problems usually have many dimensions and
  several objectives. The high dimensionality requires exceptional
  computing power, and the fact that several dimensions have to be
  searched at the same time add additional difficulty.}
% One of the most extended methods to deal with high-dimensional
% multi-objective optimization problems are the distributed
% co-evolutionary algorithms.
% ----------------- This sentence is meaningless. It should be
% substituted by other stating that co-evolutionary algorithms can
% deal with it for some reason
% Besides, we're no longer saying that they are co-evolutionary in the
% title, but "island-based selective operator application". This
% should go to the abstract too - JJ

In these methods %which ones? coevolutionary or selective operator
                   %application? - JJ
  each processor deals only with a part of the decision
space, that is, the individuals belonging to the same sub-population,
\textcolor{blue}{which are called {\em islands} in this context}, explore the same subset of decision variables. However, only
decomposable problems could be strictly implemented in parallel this
way. \textcolor{blue}{A new approach, which we call {\em selective operator application} is presented here to deal with un-decomposable
problems:} % so we are saying that high-dimensional multi-objective
          % problems are a challenge, but we are presenting a solution
          % for non-decomposable problems. What's what? - JJ
instead of sharing parts
of individuals, each processor applies the variation operators to a
specific subset of the whole individual. % this solves... - JJ
We have previously proved
that extending the length of these subsets, and therefore, sharing
\textcolor{blue}{some search space between islands, implies an improvement in different performance metrics}.
 % which results? - JJ PABLO: answered
But in this paper, a new method to automatically
adapt the size of the overlapping, or shared sections, of the
individuals has been compared to other collaborative evolutionary
techniques considering different number of islands, chromosome sizes
and benchmarks. The analysis of the obtained experimental results, by
using different metrics, shows that our approach can provide
statistically significant improvements with respect to the base
algorithm \textcolor{blue}{in high-dimensional multi-objective problems}. % in high-dimensional multi-objective problems? - JJ PABLO: ok, puesto
Results also show that the relation between the number of
islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables.  
\textcolor{blue}{This opens up a new challenge that can be addressed in the future.}
% This
                                % solves a problem or challenge we
                                % hadn't stated. - JJ Pablo: metido, mira a ver si queda frase memorable

@@

\begin{itemize}
\item Hypervolume (HV): measures the area formed by all non-dominated solutions found with respect to a reference point. Higher values imply better quality of the PF. Figure \ref{fig:hypervolume} clarifies how the hypervolume is calculated.
\item Inverted Generational distance (IGD): calculates the distance of the obtained set of solutions to the optimal PF. Therefore, this metric requires the optimal PF found in the literature, or the theoretical one. In this metric, the lower the better. %Pablo: theoretical?
\item Spread (S): Measures the spread between solutions, taking into account the \textcolor{blue}{Euclidean} distance between consecutive solutions. As in previous metric, the lower value, the better, as it implies solutions distributed along all the PF.
\end{itemize}

@@

Different metrics, previously explained, have been used
to calculate the quality of the obtained PFs in each configuration. As
some of the metrics  (such as HV) require a reference point to be
calculated, the point (1,9) has been chosen as reference, as none of
the generated PFs in all runs \textcolor{blue}{is} dominated by it. Metrics are then
normalized respecting to that point. A Kruskal-Wallis significance
test has been performed to the metrics of all runs of the
configurations, as the Kolmogorov-Smirnov test detected non-normal
distributions. The average results for each configuration are shown in
Table \ref{tab:results512} (for 512 dimensions) and Table
\ref{tab:results2048} (for 2048). As previously explained, when $P=8$
the results of A are equivalent to the obtained with D (because
$c=0$), and when $P=16$ they are equivalent to O ($c=1$). We refer this equivalence with the word {\em Equiv-} in tables. For example, Equiv-D in an HV cell  means that the HV of A is the same than the value in D column, because the behaviour of this configuration is the same than executing D.

@@

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt1_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT1 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt1_512}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt2_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT2 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt2_512}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt3_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT3 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt3_512}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt6_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT6 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt6_512}
\end{figure}

%%%%%2018

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt1_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT1 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt1_2048}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt2_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT2 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt2_2048}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt3_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT3 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt3_2048}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt6_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT6 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt6_2048}
\end{figure}

%EXPLANATION OF THE RESULTS
This can be explained comparing the number of non-dominated solutions
of each front and the average number of generations (Table
\ref{tab:sols512} and Table \ref{tab:sols2048}). With both chromosome
lengths, increasing the number of islands implies more generations
with all the methods (logically, as there are less individuals in each
island). But also, the SLA methods
approach to the number of generations with the baseline when
increasing the number of islands. However, with $L=2048$ the migration
time is big enough  to not improve the number of generations
with respect to the baseline, even in the lower number of islands, as
it happens with $L=512$. Therefore, more generations do not
necessarily improve the solution of the global PF, but focusing on
different elements of the chromosome. As  previously stated, every
island is unaware of the other islands PFs, and they are trying to
optimize their solutions independently. With respect to the average
number of solutions per front, there is a clear difference 
 with the baseline with $L=512$, where this value is in the most
of the cases, less than a half. The number of non-dominated solutions
also implies a better Spread indicator, where the baseline 
 obtains better (or not significantly different) values in the majority of the compared configurations.
